---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ucimlrepo

<!-- badges: start -->
  [![R-CMD-check](https://github.com/coatless-rpkg/ucimlrepo/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coatless-rpkg/ucimlrepo/actions/workflows/R-CMD-check.yaml)
  <!-- badges: end -->

The goal of `ucimlrepo` is to download and import data sets directly into R 
from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/).

> [!IMPORTANT]
> 
> This package is an unoffical port of the [Python `ucimlrepo` package](https://github.com/uci-ml-repo/ucimlrepo).


## Installation

You can install the development version of ucimlrepo from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("coatless-rpkg/ucimlrepo")
```

## Usage

To use `ucimlrepo`, load the package using:

```{r}
#| label: load-pkg
library(ucimlrepo)
```

With the package now loaded, we can download a dataset using the `fetch_ucirepo()` function
or use the `list_available_datasets()` function to view a list of available datasets.

### Download data

For example, to download the `iris` dataset, we can use:

```{r}
#| label: download dataset
# Fetch a dataset by name
iris_by_name <- fetch_ucirepo(name = "iris")
names(iris_by_name)
```

There are many levels to the data returned. For example, we can extract the 
original data frame containing the `iris` dataset using:


```{r}
iris_uci <- iris_by_name$data$original
head(iris_uci)
```

You can also directly query by using an ID: 

```{r}
#| label: download dataset by id
#| eval: false
# Fetch a dataset by id
iris_by_id <- fetch_ucirepo(id = 53)
```


### View list of data sets

We can also view a list of data sets available for download
using the `list_available_datasets()` function:

```{r}
#| label: list-packages
#| eval: false
# List available datasets
list_available_datasets()
```

> [!NOTE]
>
> Not all 600+ datasets on UCI ML Repo are available for download using the package. 
> The current list of available datasets can be viewed [here](https://archive.ics.uci.edu/datasets?Python=true&skip=0&take=25&sort=desc&orderBy=NumHits).
> 
> If you would like to see a specific dataset added, please submit a comment
> on an [issue ticket in the upstream repository](https://github.com/uci-ml-repo/ucimlrepo/issues/9).
